---
title: "OxFord COVID-19 Data Exploration"
output: html_document
author: "Estefania Avila, Meitong Hu, Egle Klekere, Alla Polishchuk, Jingjing Zhang"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Source & Summary

OXFORD COVID-19 GOVERNMENT RESPONSE TRACKER: aims to track and compare government responses to the coronavirus outbreak worldwide rigorously and consistently. We can analyze how specific government measures affect outcomes in terms of COVID cases and deaths.
https://www.bsg.ox.ac.uk/research/research-projects/oxford-covid-19-government-response-tracker

#### 13 Measures Tracked by Oxford:

* S1_School closing
* S2_Workplace closing
* S3_Cancel public events
* S4_Close public transport
* S5_Public information campaigns
* S6_Restrictions on internal movement
* S7_International travel controls
* S8_Fiscal measures
* S9_Monetary measures
* S10_Emergency investment in health care
* S11_Investment in Vaccines
* S12_Testing framework
* S13_Contact tracing

OxCGRT collects publicly available information on 13 indicators of government response (S1-S13). Nine of the indicators (S1-S7, S12 and S13) take policies such as school closures, travel bans, etc. are recorded on an ordinal scale; the remainder (S8-S11) are financial indicators such as fiscal or monetary measures.

#### Other data in dataset:

* ConfirmedCases
* ConfirmedDeaths
* StringencyIndex

Oxford has a StringencyIndex that measures how stringent each government has reacted. I thought, maybe we can introduce Proactivity index - how many measures country has introduced before first ConfirmedCase and first ConfirmedDeath. I assume this affects the outcomes as well. Then one of our questiond would be: What has been more effective stringency or proactivity?

Interesting analysis of Oxford data:
https://www.bsg.ox.ac.uk/research/research-projects/oxford-covid-19-government-response-tracker?ref=researchstash

## Data Exploration
### Read File
Read file in.Noticed it has a lot of "Notes" columns used to supply source of change in variable.

```{r ,}
data = read.csv("OxCGRT_Download_080420_161822_Full.csv")
names(data)
```

### Filtered DataFrame
We removed those columns to facilitate initial data exploration.

```{r ,}
df = subset( data, select = -c(grep( "Notes" , names( data ) )) )
names(df)
str(df)
summary(df)

```

### Explore Notes
To explore the notes data, we just peaked into one of the columns to see what kind of announcements they are using.
```{r ,}
df_notes = subset( data, select = c(grep( "Notes" , names( data ) )) )
names(df_notes)
str(df_notes)
levels(df_notes$S1_Notes)[1:5]
```
