---
title: "OxFord COVID-19 Data Exploration"
output: html_document
author: "Estefania Avila, Meitong Hu, Egle Klekere, Alla Polishchuk, Jingjing Zhang"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Source

OXFORD COVID-19 GOVERNMENT RESPONSE TRACKER: aims to track and compare government responses to the coronavirus outbreak worldwide rigorously and consistently. We can analyze how specific government measures affect outcomes in terms of COVID cases and deaths.
https://www.bsg.ox.ac.uk/research/research-projects/oxford-covid-19-government-response-tracker

13 Measures Tracked by Oxford:

S1_School closing

S2_Workplace closing

S3_Cancel public events

S4_Close public transport

S5_Public information campaigns

S6_Restrictions on internal movement

S7_International travel controls

S8_Fiscal measures

S9_Monetary measures

S10_Emergency investment in health care

S11_Investment in Vaccines

S12_Testing framework

S13_Contact tracing

ConfirmedCases

ConfirmedDeaths

StringencyIndex

Interesting analysis of Oxford data:
https://www.bsg.ox.ac.uk/research/research-projects/oxford-covid-19-government-response-tracker?ref=researchstash

Oxford has a StringencyIndex that measures how stringent each government has reacted. I thought, maybe we can introduce Proactivity index - how many measures country has introduced before first ConfirmedCase and first ConfirmedDeath. I assume this affects the outcomes as well. Then one of our questiond would be: What has been more effective stringency or proactivity?

## Data Explore
### Read File
Read file in.Noticed it has a lot of "Notes" columns used to supply source of change in variable. So we removed those columns to facilitate initial data exploration.

```{r ,}
data = read.csv("OxCGRT_Download_080420_161822_Full.csv")
names(data)
```

### Filtered DataFrame
Noticed the data has a lot of "Notes" columns used to supply source of change in variable. So we removed those columns to facilitate initial data exploration.

```{r ,}
df = subset( data, select = -c(grep( "Notes" , names( data ) )) )
names(df)
str(df)
summary(df)

```

### Explore Notes
To explore the notes data, we just peaked into one of the columns to see what kind of announcements they are using.
```{r ,}
df_notes = subset( data, select = c(grep( "Notes" , names( data ) )) )
names(df_notes)
str(df_notes)
levels(df_notes$S1_Notes)
```
